Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-12-11 text/ms.tex
Tweak text for thesis formatting, see notes
- Supporting Information -> Supporting materials
- Figures weaved throughout
- Embed som.tex at end

We selected abundance time series from the Global Population Dynamics Database
(GPDD; \citeauthor{gpdd2010} \citeyear{gpdd2010}), which contains nearly 5000
time series of abundance from $\sim$1000 species and $\sim$100 taxonomic
orders. We filtered the data (Supporting Information) to remove populations
from less reliable data sources, and those without sufficient data for our
models, and then interpolated some missing values
\citep[\textit{sensu}][]{brook2006a}. Our interpolation affected only
$\sim$\interpPointsPerc \% of the final data points (Table~\ref{tab:stats})
and none of the data points that were later considered black-swan events. Our
final dataset contained \NPops~populations across \NOrders~taxonomic orders
and seven taxonomic classes, with a median of \medianTimeSteps~time steps
(range of \minTimeSteps--\maxTimeSteps) (Table~\ref{tab:stats},
Fig.~\ref{fig:all-ts}).

@@

We fit all models in a Bayesian framework using Stan \citep{stan-manual2014}
via the R computing environment \citep{r2014}. Stan samples from the posterior
distribution with an adaptive version of Hamiltonian Markov chain Monte Carlo
called the No-U-Turn Sampler and generally obtains less correlated samples
than algorithms such as the Gibbs sampler \citep{hoffman2014}. We tested to
ensure that the chains had sufficiently converged and that the sampler had
obtained sufficient independent samples from the posterior ($\widehat{R} <
1.05$, $n_\mathrm{eff} > 200$; Supporting Information).

@@

We fit alternative population models to test if four key phenomena
systematically changed our conclusions. Autocorrelation has been suggested as
a reason for increased observed variability of abundance time series through
time, which could create apparent heavy tails \citep{inchausti2002};
therefore, we fit a model that included serial correlation in the residuals.
Additionally, previous work has modelled abundance or growth rates without
accounting for density dependence \citep{halley2002,segura2013}; therefore, we
fit a simpler model in which we assumed density independence. Third,
observation error could bias parameter estimates \citep{knape2012} or mask our
ability to detect heavy tails \citep{ward2007}; therefore, we fit a model
where we allowed for a fixed quantity of observation error ($0.2$ standard
deviations on a log scale). Finally, the Gompertz model assumes that
population growth rate declines linearly with log abundance. Therefore, we
also fit an alternative model, the Ricker-logistic model, which assumes that
population growth rate declines linearly with abundance itself (Supporting
Information).

In addition to alternative population models, we investigated the sensitivity
of our results to weaker and stronger priors (exponential rate parameter $=
0.005, 0.02$; Fig.~\ref{fig:priors}, Supporting Information). We used
simulated data to test how easily we could detect $\nu$ given different sample
sizes and to ensure we could recover unbiased parameter estimates from the
Gompertz model (Supporting Information).

@@

We investigated possible covariates of heavy-tailed population dynamics
visually and through multilevel modelling. We plotted characteristics of the
time series ($\sigma$, $\lambda$, $b$, and time-series length) along with two
life-history characteristics (body length and maximum lifespan obtained from
\citet{brook2006a}) against our estimated probability of heavy tails, Pr$(\nu <
10)$. We formally investigated these relationships by fitting beta
regression multilevel models \citep{ferrari2004}. The beta probability
distribution can represent continuous response data that range between zero
and one; we fit our models with a logit link as is common for beta and logistic
regression \citep{ferrari2004}. We incorporated standard deviations around the
means for covariates that were derived from Gompertz model parameter
estimates. To account for broad patterns of phylogenetic relatedness, we
allowed for hierarchical intercepts at the taxonomic class, order, and species
level (Supporting Information). We fit our model in Stan with weakly
informative priors on the coefficients \citep{gelman2008d} and variance
parameters (\citeauthor{gelman2006c} \citeyear{gelman2006c}
\citeauthor{gelman2014} \citeyear{gelman2014}; Supporting Information).

@@

We found strong, but rare, evidence for black-swan population dynamics. By
defining black-swan dynamics as a greater than $0.5$ probability that $\nu <
10$, our main Gompertz model found evidence for heavy tails most frequently
for birds (\birdPH\%) followed by mammals (\mammalsPH\%), and insects
(\insectsPH\%) (Fig.~\ref{fig:nu-coefs}, Table~\ref{tab:causes-supp}). Black
swans were taxonomically widespread, occurring in \POrdersHeavy\% of taxonomic
orders. Accounting for time series length and partially pooling inference
across taxonomic class and order with a multilevel model (Supporting
Information), there was stronger evidence for black swans in insect
populations than is visually apparent in Fig.~\ref{fig:nu-coefs}---four of 10
orders with the highest median probability of heavy tails were insect
orders---however, there was considerable uncertainty in these estimates
(Fig.~\ref{fig:posteriors}a).

The majority of our heavy-tailed estimates were robust to alternative
population models, observation error, and choice of priors. Our conclusions
were not systematically altered when we included an autocorrelation structure
in the residuals, modelled population growth rates without density dependence,
or modelled the population dynamics as Ricker-logistic (Fig.~\ref{fig:alt}).
However, setting observation error standard deviation to $0.2$ increased the
median estimate of $\nu$ from $<10$ to $\ge 10$ in $\baseNuTenObsTenSwitch$ of
$\baseNuTen$ populations, although the majority of $\nu$ estimates remained
qualitatively similar (Fig.~\ref{fig:alt}). The strength of the prior on $\nu$
had little influence on estimates of black-swan dynamics
(Fig.~\ref{fig:alt-priors}). Our simulation testing shows that, if anything,
our models underpredict the true magnitude and probability of heavy tailed
events---especially given the length of the time series in the GPDD
(Figs~\ref{fig:sim-nu}, \ref{fig:sim-prob}).

@@

We examined all time series with published explanations of why the black-swan
events occurred (Tables~\ref{tab:sparks} and \ref{tab:causes-supp}). The
majority of documented events (\percBSDown \%) were downward black swans and
involved a combination of multiple factors. For example, a synchronization of
environmental- and predation-mediated population cycles are thought to have
caused a downward black-swan event for a water vole (\textit{Arvicola
  terrestris}) population \citep{saucy1994}. Other black swans were the result
of a sequence of extreme climate events on their own. For instance, severe
winters in 1929, 1940--1942, and 1962--1963 were associated with black-swan
downswings in grey heron (\textit{Ardea cinerea}) abundance in the United
Kingdom \citep{stafford1971}. Our analysis finds that the last event was a
combination of two black-swan events in a row and it took the population three
times longer to recover than predicted \citep{stafford1971}. Downwards black
swans were sometimes followed by upwards black swans. For example, during a
period of population crowding and nest shortages, a population of European
shag cormorants (\textit{Phalacrocorax aristotelis}) on the Farne Islands,
United Kingdom, declined suddenly following a red tide event in 1968
\citep{potts1980}. This freed quality nest sites for first-time breeders,
productivity rapidly increased, and the population experienced a rapid upswing
in abundance. \citep{potts1980}.

@@

There are a number of caveats when considering the generality of our results.
The GPDD data represent a taxonomically and geographically biased sample of
populations---the longer time series we focus on are dominated by commercially
and recreationally important species and a disproportionate number of
populations are located in the United Kingdom. Although we would expect to
find qualitatively similar evidence for black swans in many other large
taxonomic or geographic samples of populations, the common forces driving
those black swans would likely differ (e.g.~severe winters in
Table~\ref{tab:sparks}). In addition to a possibly biased sample of
populations, some black-swan detections could just be recording mistakes, or
conversely, some extreme observations may have been discarded or altered if
they were erroneously suspected of being recording mistakes. Indeed, we
discarded three of the populations that our method initially identified as
heavy tailed because they turned out to be data-entry errors (Supporting
Information). A final caveat is that the temporal scales of observation and
population dynamics vary considerably across populations in the GPDD and this
likely influences the detection of heavy tails. As an example, if we make
frequent observations relative to generation time (e.g.~for many large-bodied
mammals) we will average across generations and perhaps miss black swans.
Conversely, if we census populations infrequently relative to generation time
(e.g.~many insects in the GPDD) the recorded data may average across extreme
and less-extreme events and also dampen black-swan dynamics.

@@

We thank J.W. Moore, J.D. Yeakel, and other members of the Earth to Ocean
research group for helpful discussions and comments. We are grateful to the
contributors and maintainers of the Global Population Dynamics Database and to
Compute Canada's WestGrid high-performance computing resources. Silhouette
images were obtained from \texttt{phylopic.org} under Creative Commons
licenses; sources are listed in the Supporting Information. Funding was
provided by a Simon Fraser University Graduate Fellowship (SCA), the Natural
Sciences and Engineering Research Council of Canada (NKD, ABC), the Canada
Research Chairs Program (NKD).

\section{Supporting Information}

The following supporting information is available online for this article:\\
Tables S1 and S2\\
Figures S1--S6\\
Example Stan code for a heavy-tailed Gompertz model and multilevel beta
regression\\
The GPDD IDs used in our analysis\\
All code and data to recreate our analysis are available at:\\
\url{https://github.com/seananderson/heavy-tails}

\renewcommand{\baselinestretch}{\tighttextstretch}
\normalsize
\bibliographystyle{apalike}
\bibliography{/Users/seananderson/Dropbox/tex/jshort,/Users/seananderson/Dropbox/tex/ref3}

\clearpage
\renewcommand{\baselinestretch}{\textstretch}
\normalsize

\section{Tables}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/t-nu-eg2.pdf}
\caption[An illustration of fitting population dynamic models that allow for heavy
tails, represented by the Student-t degrees of freedom parameter $\nu$.]{
An illustration of fitting population dynamic models that allow for heavy
tails, represented by the Student-t degrees of freedom parameter $\nu$. (a, b)
The probability density for t distributions with a scale parameter of 1 and
different values of $\nu$. Small values of $\nu$ create heavy tails. As $\nu$
approaches infinity the distribution approaches the normal distribution. For
example, at $\nu = 2$, the probability of drawing a value more than five
standard deviations below the mean is 1.8\%, whereas the probability of drawing
such a value from a normal distribution is nearly zero ($2.9\cdot10^{-5}$\%).
(c--e) Simulated population dynamics from a Gompertz model with process noise
drawn from t distributions with three different values of $\nu$. Coloured dots
in panels c and d represent jumps with less than a 1 in 1000 chance of
occurring in a normal distribution. (f--h) Estimates of $\nu$ from models fit
to the times series in panels c--e. Shown are the posterior samples
(histograms), median and interquartile range of the posterior (IQR) (dots and
line segments), and the exponential prior on $\nu$ (dashed lines). Colour
shading behind panels f--h illustrates the region of heavy tails.
}
\label{fig:didactic}
\end{center}
\end{figure}

\clearpage

\label{fig:nu-coefs}
\end{center}
\end{figure}
\clearpage

\caption[Posterior probability distributions from beta regression multilevel
  models.]{Posterior probability distributions from beta regression multilevel
  models. (a) Taxonomic-order-level posterior densities of Pr($\nu < 10$)
  (approximately the probability of heavy tails) after accounting for
  time-series length. Estimates are at the geometric mean of time series
  length across all the data (approximately 27 time steps). Colour shading
  refers to taxonomic class (yellow: fishes, green: insects, purple: birds,
  and red: mammals). Dotted vertical line in panel a indicates the median
  expected Pr($\nu < 10$) from the prior distribution. (b) Main effect
  posterior densities for potential covariates of Pr($\nu < 10$). The beta
  regression models were fit on a logit scale with hierarchical intercepts for
  taxonomic class, taxonomic order, and species. All covariates were
  standardized by subtracting their mean and dividing by twice their standard
  deviation. In both panels, short vertical line segments within the density
  polygons indicate median posterior estimates.}

\caption[Potential covariates of heavy-tailed population dynamics (indicated
  by a high probability that $\nu < 10$).]{Potential covariates of heavy-tailed population dynamics (indicated
  by a high probability that $\nu < 10$). Shown are (a--c) parameters from the
  Gompertz heavy-tailed population model ($\sigma$, $\lambda$, $b$), (b)
  number of time steps, (c) body length, and (d) lifespan. For the Gompertz
  parameters, $\sigma$ refers to the scale parameter of the Student-t
  process-noise distribution, $\lambda$ refers to the expected log abundance
  at the next time step at an abundance of one, $b$ refers to the density
  dependence parameter ($1$ is maximally density independent, $0$ is maximally
  density dependent, and $<0$ is inversely density dependent). Circles
  representing a few sharks, crustaceans, and gastropods are filled in white.
  Median and 90\% credible interval posterior predictions of a beta regression
  multilevel model are shown in panels a and d where there was a high
  probability the slope coefficient was different from zero
  (Fig.~\ref{fig:posteriors}b).}

>>>>

We selected abundance time series from the Global Population Dynamics Database
(GPDD; \citeauthor{gpdd2010} \citeyear{gpdd2010}), which contains nearly 5000
time series of abundance from $\sim$1000 species and $\sim$100 taxonomic
orders. We filtered the data (Supporting materials) to remove populations
from less reliable data sources, and those without sufficient data for our
models, and then interpolated some missing values
\citep[\textit{sensu}][]{brook2006a}. Our interpolation affected only
$\sim$\interpPointsPerc \% of the final data points (Supporting materials~Table~\ref{tab:stats})
and none of the data points that were later considered black-swan events. Our
final dataset contained \NPops~populations across \NOrders~taxonomic orders
and seven taxonomic classes, with a median of \medianTimeSteps~time steps
(range of \minTimeSteps--\maxTimeSteps) (Supporting materials Table~\ref{tab:stats},
Fig.~\ref{fig:all-ts}).

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/t-nu-eg2.pdf}
\caption[An illustration of fitting population dynamic models that allow for heavy
tails, represented by the Student-t degrees of freedom parameter $\nu$.]{
An illustration of fitting population dynamic models that allow for heavy
tails, represented by the Student-t degrees of freedom parameter $\nu$. (a, b)
The probability density for t distributions with a scale parameter of 1 and
different values of $\nu$. Small values of $\nu$ create heavy tails. As $\nu$
approaches infinity the distribution approaches the normal distribution. For
example, at $\nu = 2$, the probability of drawing a value more than five
standard deviations below the mean is $0.02$, whereas the probability of drawing
such a value from a normal distribution is nearly zero ($2.9\cdot10^{-7}$).
(c--e) Simulated population dynamics from a Gompertz model with process noise
drawn from t distributions with three different values of $\nu$. Coloured dots
in panels c and d represent jumps with less than a 1 in 1000 chance of
occurring in a normal distribution. (f--h) Estimates of $\nu$ from models fit
to the times series in panels c--e. Shown are the posterior samples
(histograms), median and interquartile range of the posterior (IQR) (dots and
line segments), and the exponential prior on $\nu$ (dashed lines). Colour
shading behind panels f--h illustrates the region of heavy tails.
}
\label{fig:didactic}
\end{center}
\end{figure}

We fit all models in a Bayesian framework using Stan \citep{stan-manual2014}
via the R computing environment \citep{r2014}. Stan samples from the posterior
distribution with an adaptive version of Hamiltonian Markov chain Monte Carlo
called the No-U-Turn Sampler and generally obtains less correlated samples
than algorithms such as the Gibbs sampler \citep{hoffman2014}. We tested to
ensure that the chains had sufficiently converged and that the sampler had
obtained sufficient independent samples from the posterior ($\widehat{R} <
1.05$, $n_\mathrm{eff} > 200$; Supporting materials).

@@

We fit alternative population models to test if four key phenomena
systematically changed our conclusions. Autocorrelation has been suggested as
a reason for increased observed variability of abundance time series through
time, which could create apparent heavy tails \citep{inchausti2002};
therefore, we fit a model that included serial correlation in the residuals.
Additionally, previous work has modelled abundance or growth rates without
accounting for density dependence \citep{halley2002,segura2013}; therefore, we
fit a simpler model in which we assumed density independence. Third,
observation error could bias parameter estimates \citep{knape2012} or mask our
ability to detect heavy tails \citep{ward2007}; therefore, we fit a model
where we allowed for a fixed quantity of observation error ($0.2$ standard
deviations on a log scale). Finally, the Gompertz model assumes that
population growth rate declines linearly with log abundance. Therefore, we
also fit an alternative model, the Ricker-logistic model, which assumes that
population growth rate declines linearly with abundance itself (Supporting
materials).

In addition to alternative population models, we investigated the sensitivity
of our results to weaker and stronger priors (exponential rate parameter $=
0.005, 0.02$; Supporting materials Fig.~\ref{fig:priors}, Supporting materials). We used
simulated data to test how easily we could detect $\nu$ given different sample
sizes and to ensure we could recover unbiased parameter estimates from the
Gompertz model (Supporting materials).

@@

We investigated possible covariates of heavy-tailed population dynamics
visually and through multilevel modelling. We plotted characteristics of the
time series ($\sigma$, $\lambda$, $b$, and time-series length) along with two
life-history characteristics (body length and maximum lifespan obtained from
\citet{brook2006a}) against our estimated probability of heavy tails, Pr$(\nu <
10)$. We formally investigated these relationships by fitting beta
regression multilevel models \citep{ferrari2004}. The beta probability
distribution can represent continuous response data that range between zero
and one; we fit our models with a logit link as is common for beta and logistic
regression \citep{ferrari2004}. We incorporated standard deviations around the
means for covariates that were derived from Gompertz model parameter
estimates. To account for broad patterns of phylogenetic relatedness, we
allowed for hierarchical intercepts at the taxonomic class, order, and species
level (Supporting materials). We fit our model in Stan with weakly
informative priors on the coefficients \citep{gelman2008d} and variance
parameters (\citeauthor{gelman2006c} \citeyear{gelman2006c};
\citeauthor{gelman2014} \citeyear{gelman2014}; Supporting materials).

@@

We found strong, but rare, evidence for black-swan population dynamics. By
defining black-swan dynamics as a greater than $0.5$ probability that $\nu <
10$, our main Gompertz model found evidence for heavy tails most frequently
for birds (\birdPH\%) followed by mammals (\mammalsPH\%), and insects
(\insectsPH\%) (Fig.~\ref{fig:nu-coefs}, Supporting materials Table~\ref{tab:causes-supp}). Black
swans were taxonomically widespread, occurring in \POrdersHeavy\% of taxonomic
orders. Accounting for time series length and partially pooling inference
across taxonomic class and order with a multilevel model (Supporting
materials), there was stronger evidence for black swans in insect
populations than is visually apparent in Fig.~\ref{fig:nu-coefs}---four of 10
orders with the highest median probability of heavy tails were insect
orders---however, there was considerable uncertainty in these estimates
(Fig.~\ref{fig:posteriors}a).

\label{fig:nu-coefs}
\end{center}
\end{figure}

\caption[Posterior probability distributions from beta regression multilevel
  models.]{Posterior probability distributions from beta regression multilevel
  models. (a) Taxonomic-order-level posterior densities of Pr($\nu < 10$)
  (approximately the probability of heavy tails) after accounting for
  time-series length. Estimates are at the geometric mean of time series
  length across all the data (approximately 27 time steps). Colour shading
  refers to taxonomic class (yellow: fishes, green: insects, purple: birds,
  and red: mammals). Dotted vertical line in panel a indicates the
  Pr($\nu < 10$) from the prior distribution. (b) Main effect
  posterior densities for potential covariates of Pr($\nu < 10$). The beta
  regression models were fit on a logit scale with hierarchical intercepts for
  taxonomic class, taxonomic order, and species. All covariates were
  standardized by subtracting their mean and dividing by twice their standard
  deviation. In both panels, short vertical line segments within the density
  polygons indicate median posterior estimates.}

The majority of our heavy-tailed estimates were robust to alternative
population models, observation error, and choice of priors. Our conclusions
were not systematically altered when we included an autocorrelation structure
in the residuals, modelled population growth rates without density dependence,
or modelled the population dynamics as Ricker-logistic (Supporting materials Fig.~\ref{fig:alt}).
However, setting observation error standard deviation to $0.2$ increased the
median estimate of $\nu$ from $<10$ to $\ge 10$ in $\baseNuTenObsTenSwitch$ of
$\baseNuTen$ populations, although the majority of $\nu$ estimates remained
qualitatively similar (Supporting materials Fig.~\ref{fig:alt}). The strength of the prior on $\nu$
had little influence on estimates of black-swan dynamics
(Supporting materials~Fig.~\ref{fig:alt-priors}). Our simulation testing shows that, if anything,
our models underpredict the true magnitude and probability of heavy tailed
events---especially given the length of the time series in the GPDD
(Supporting materials Figs~\ref{fig:sim-nu}, \ref{fig:sim-prob}).

@@

\caption[Potential covariates of heavy-tailed population dynamics.]{Potential covariates of heavy-tailed population dynamics (indicated
  by a high probability that $\nu < 10$). Shown are (a--c) parameters from the
  Gompertz heavy-tailed population model ($\sigma$, $\lambda$, $b$), (b)
  number of time steps, (c) body length, and (d) lifespan. For the Gompertz
  parameters, $\sigma$ refers to the scale parameter of the Student-t
  process-noise distribution, $\lambda$ refers to the expected log abundance
  at the next time step at an abundance of one, $b$ refers to the density
  dependence parameter ($1$ is maximally density independent, $0$ is maximally
  density dependent, and $<0$ is inversely density dependent). Circles
  representing a few sharks, crustaceans, and gastropods are filled in white.
  Median and 90\% credible interval posterior predictions of a beta regression
  multilevel model are shown in panels a and d where there was a high
  probability the slope coefficient was different from zero
  (Fig.~\ref{fig:posteriors}b).}

We examined all time series with published explanations of why the black-swan
events occurred (Table~\ref{tab:sparks} and Supporting materials~Table~\ref{tab:causes-supp}). The
majority of documented events (\percBSDown \%) were downward black swans and
involved a combination of multiple factors. For example, a synchronization of
environmental- and predation-mediated population cycles are thought to have
caused a downward black-swan event for a water vole (\textit{Arvicola
  terrestris}) population \citep{saucy1994}. Other black swans were the result
of a sequence of extreme climate events on their own. For instance, severe
winters in 1929, 1940--1942, and 1962--1963 were associated with black-swan
downswings in grey heron (\textit{Ardea cinerea}) abundance in the United
Kingdom \citep{stafford1971}. Our analysis finds that the last event was a
combination of two black-swan events in a row and it took the population three
times longer to recover than predicted \citep{stafford1971}. Downwards black
swans were sometimes followed by upwards black swans. For example, during a
period of population crowding and nest shortages, a population of European
shag cormorants (\textit{Phalacrocorax aristotelis}) on the Farne Islands,
United Kingdom, declined suddenly following a red tide event in 1968
\citep{potts1980}. This freed quality nest sites for first-time breeders,
productivity rapidly increased, and the population experienced a rapid upswing
in abundance \citep{potts1980}.

@@

There are a number of caveats when considering the generality of our results.
The GPDD data represent a taxonomically and geographically biased sample of
populations---the longer time series we focus on are dominated by commercially
and recreationally important species and a disproportionate number of
populations are located in the United Kingdom. Although we would expect to
find qualitatively similar evidence for black swans in many other large
taxonomic or geographic samples of populations, the common forces driving
those black swans would likely differ (e.g.~severe winters in
Table~\ref{tab:sparks}). In addition to a possibly biased sample of
populations, some black-swan detections could just be recording mistakes, or
conversely, some extreme observations may have been discarded or altered if
they were erroneously suspected of being recording mistakes. Indeed, we
discarded three of the populations that our method initially identified as
heavy tailed because they turned out to be data-entry errors (Supporting
materials). A final caveat is that the temporal scales of observation and
population dynamics vary considerably across populations in the GPDD and this
likely influences the detection of heavy tails. As an example, if we make
frequent observations relative to generation time (e.g.~for many large-bodied
mammals) we will average across generations and perhaps miss black swans.
Conversely, if we census populations infrequently relative to generation time
(e.g.~many insects in the GPDD) the recorded data may average across extreme
and less-extreme events and also dampen black-swan dynamics.

@@

We thank J.W. Moore, J.D. Yeakel, and other members of the Earth to Ocean
research group for helpful discussions and comments. We are grateful to the
contributors and maintainers of the Global Population Dynamics Database and to
Compute Canada's WestGrid high-performance computing resources. Silhouette
images were obtained from \texttt{phylopic.org} under Creative Commons
licenses; sources are listed in the Supporting materials. Funding was provided
by a Simon Fraser University Graduate Fellowship (SCA), the Natural Sciences
and Engineering Research Council of Canada (NKD, ABC), the Canada Research
Chairs Program (NKD).

% \section{Supporting Information}
%
% The following supporting information is available online for this article:\\
% Tables S1 and S2\\
% Figures S1--S6\\
% Example Stan code for a heavy-tailed Gompertz model and multilevel beta
% regression\\
% The GPDD IDs used in our analysis\\
% All code and data to recreate our analysis are available at:\\
% \url{https://github.com/seananderson/heavy-tails}
%
% \renewcommand{\baselinestretch}{\tighttextstretch}
% \normalsize
% \bibliographystyle{apalike}
% \bibliography{/Users/seananderson/Dropbox/tex/jshort,/Users/seananderson/Dropbox/tex/ref3}
%
% \clearpage
% \renewcommand{\baselinestretch}{\textstretch}
% \normalsize

%\section{Tables}

%\clearpage

%\clearpage

%\clearpage

%\clearpage
\input{blackswans/som}
